---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Triangulate

<!-- badges: start -->
<!-- badges: end -->

This is a small R package that provides functions for triangulating five points into three triangles. It includes a triangulation function, a plotting utility and example datasets

## Installation

You can install the development version of Triangulate from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("gkagyen/Triangulate")
```
## Features
**`triangulate()`**: This function connects five points (given a 5x2 matrix) into three triangles returning a list with the points and edges. `triangulate()` connects the five points using a fixed pattern: 1-2-3, 2-4-5, 3-5-2.

**`plot_triangles()`**: This function helps to visualise the triangulation using `ggplot2`, labelling points with their respective coordinates.

**Example datasets**: Two built-in 5x2 matrices (`sample_data1` and `sample_data2`) for demonstration

## Example Usage

Here is a quick example to triangulate and plot five points

```{r example}
library(Triangulate)

# create a 5x2 matrix of points
points <- matrix(c(0,0,1,1,2,0,3,1,4,0), nrow = 5, byrow = TRUE)

# apply triangulation
tri_obj <- triangulate(points)

# visualise triangulation
tri_show <- plot_triangles(tri_obj)
print(tri_show)
```

### Using In-Built Datasets
```{r}
sample_data1
plot_triangles(triangulate(sample_data1))

sample_data2
plot_triangles(triangulate(sample_data2))
```

## Motivations
This package was developed as part of the ***Google Summer of Code 2025*** application under the R projects' **"Create an R Code only version of the `tripack` functions"** project
